Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-05-08 paper/AlignedIsotropicGW.tex
Realised arXiver will only upload take three figures in a comment.
Add section on param-dependent chi_eff constraints

% arXiver will snag the right plots so they show up on Twitter :)
%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,chi-eff-mock-posteriors.pdf,six-way-O2-model-selection.pdf}

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two years
\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
additional detections drawn from each of our six zero-parameter models
for the spin distribution in addition to the three existing detections
from O1, with observational uncertainties drawn randomly from the
three Gaussian widths used to approximate the $\chieff$ posteriors in
Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
  $\chieff$ depends on the other parameters of the merging \ac{BBH}
  system, particularly on the mass ratio.  Our assumption about future
  observational uncertainties is appropriate if the parameters of the
  three detected events are representative of the parameters of future
  detections.}, we find the odds ratios shown in Figure
\ref{fig:O2-predictions}.  We find that most scenarios with an
additional 10 detections allow the simulated angular distribution to
be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
credibility; and in the most pessimistic case the distinction is
typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
odds ratio).  While such future detections should permit a confident
distinction between \emph{angular} distributions, we would remain much
less certain about the \emph{magnitude} distribution among the three
options considered here until we have a larger number of observations.
In Figure \ref{fig:O2-predictions}, the odds ratio between different
magnitude distributions with the same angular distribution is much
closer to unity than the odds ratio between angular distributions.

@@

>>>>

% arXiver will snag the right plots so they show up on Twitter :)
%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,six-way-O2-model-selection.pdf}

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two years
\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
additional detections drawn from each of our six zero-parameter models
for the spin distribution in addition to the three existing detections
from O1, with observational uncertainties drawn randomly from the
three Gaussian widths used to approximate the $\chieff$ posteriors in
Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
  $\chieff$ depends on the other parameters of the merging \ac{BBH}
  system, particularly on the mass ratio.  Our assumption about future
  observational uncertainties is appropriate if the parameters of the
  three detected events are representative of the parameters of future
  detections. See Section \ref{sec:chi-eff-precision}
for further discussion.}, we find the odds ratios shown in Figure
\ref{fig:O2-predictions}.  We find that most scenarios with an
additional 10 detections allow the simulated angular distribution to
be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
credibility; and in the most pessimistic case the distinction is
typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
odds ratio).  While such future detections should permit a confident
distinction between \emph{angular} distributions, we would remain much
less certain about the \emph{magnitude} distribution among the three
options considered here until we have a larger number of observations.
In Figure \ref{fig:O2-predictions}, the odds ratio between different
magnitude distributions with the same angular distribution is much
closer to unity than the odds ratio between angular distributions.

@@

\section{Precision of $\chieff$ measurements}
\label{sec:chi-eff-precision}
Throughout this work we have made the simplifying assumption that the
precision to which $\chieff$ can be constrained for individual binaries
is independent of the binary's properties.  In practice, our ability to
constrain $\chieff$ \emph{is} dependent on the system's properties, in
particular its true $\chieff$ and mass ratio, which we illustrate in
figure \ref{fig:chi-eff-constraints}.

For this figure a detected population\footnote{We qualify a system as
``detected'' if it produces a SNR above $8$ in the second-loudest
detector to select only coincident events.} of 500 binaries was selected
from a population with component masses distributed uniformly between
$1$ and $30~\mathrm{M}_\odot$ with $m_1 + m_2 < 30~\mathrm{M}_\odot$,
locations distributed uniformly in volume, and orientations distributed
isotropically.  Data were simulated for each binary, and posteriors were
estimated using the LIGO-Virgo parameter estimation library
\texttt{LALInference}~\citep{2015PhRvD..91d2003V} using inspiral-only
waveform models (merger and ringdown effects can provide additional
information for some binaries, but we ignore those effects here).
$\chieff$ is better constrained for binaries with low (negative)
effective spins and low mass ratios.

By neglecting these effects, We do not expect to qualitatively affect
out conclusions, though they could affect predictions for the total
number of detections necessary to constrain the population.  For
example, if the universe preferentially forms asymmetric binaries with
low mass ratios, individual $\chieff$ constraints will be systematically
better, reducing the number of binaries needed to infer the properties
of the population.

\begin{figure}
  \plotone{../plots/chi-eff-90cl}
  \caption{Widths of the 90\% credible intervals for $\chieff$ for 500
    binaries in a simulated detected population. $\chieff$ is better
    constrained for systems with low $\chieff$ and low mass ratio.}
  \label{fig:chi-eff-constraints}
\end{figure}
